Handling String-Number Conversion in String Constraint Solving

### 2019-11-16 draft_saturday.tex
struggle struggle struggle, bloody lemmas bloody lemmas bloody lemmas

\newcommand{\nat}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
%\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\mathbb{P}}(#1)} %shouldn't it be something that takes automaton as parameter and returns formula?
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}
\newcommand{\encode}[1]{\mathit{encode}_{#1}}
\newcommand{\eqwrt}[1]{=_{#1}}
\newcommand{\iequiv}{\equiv}
\newcommand{\restrict}[2]{#1{\restriction_{#2}}}
\newcommand{\underf}[2]{\mathit{under}_#1(#2)}

@@

We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the sets of natural numbers (including 0) and 
integers. For a set $A$, we use $|A|$ to denote its size. 
For $n,m\in\nat$, we write $[n,m]$ for the set of natural numbers 
%(or symbols in $\Sigma$) 
$\{k\mid n\leq k \leq m\}$. 
The function $f$ with the domain restricted to $D$ is denoted by $\restrict f D$,
and a set of functions $F$ restricted to $D$ is $\restrict F D = \{\restrict f D \mid f \in F\}$.
An \emph{alphabet} is a finite set $\Sigma$ of \emph{characters} and a \emph{word} over $\Sigma$ is a sequence $w = a_1\ldots a_n$ of characters from $\Sigma$, with $\epsilon$ denoting the \emph{empty word}. 
We use $w_1\cdot w_2$ to denote the \emph{concatenation} of words $w_1$ and $w_2$.
$\Sigma^*$ is the set of all words over $\Sigma$, $\Sigma^+ = \Sigma^*\setminus \{\epsilon\}$, $\Sigma_\epsilon = \Sigma\cup\{\epsilon\}$, and 
A \emph{language} over $\Sigma$ is a subset $L$ of $\Sigma^*$.  
%
We use $|w|$ to denote the length of $w$ and $|w|_a$ to denote the number of occurrences of the character $a\in \Sigma$ in $w$.

@@

%Through the paper, we will use quantifier-free linear integer arithmetic formulas, and call them \emph{linear formulas} for short.   
%Given a linear formula $\phi$, a set of variables $V$, and an \emph{interpretation} of $V$, i.e., a function $I:V\rightarrow \nat$, 
%we denote by $I\models_V \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a \emph{model} of $\phi$ over $V$. 
%
%\lh{Our use of indexing by $V$ in the text later is not very nice. Add this for clarity?: If $V$ contain additional variables not present in $\phi$, the these might be assigned arbitrary values by $I$. If $V$ is missing some of the variables of $\phi$, then it must be possible to complete $I$ with assignment of the missing variables into a satisfying assignment of $\phi$ (as in existential quantification).}
%We use $\modelsof \phi$ to denote the set all models of $\phi$ over $V$. 
%We drop the reference to $V$ and write just ``model of $\phi$'',  $\modelsof \phi$, and $\models$ when no confusion may arise. For a set of interpretations $S$, we use $S_\Sigma$ to denote the set of interpretations obtained from $S$ by projecting the domain to $\Sigma$.

Through the paper, we will use quantifier-free linear integer arithmetic formulas, and call them \emph{linear formulas} for short.   
Given a linear formula $\phi$ over variables $V$ and an \emph{integer interpretation} of $V$, a function $I:V\rightarrow \integers$, 
we denote by $I\models \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a \emph{model} of $\phi$. 
We use $\modelsof \phi$ to denote the set all models of $\phi$. 
%
%Let $J$ and $J'$ be two sets integer interpretations, $J$ containting interpretation of variables $V$ and $J'$ of variables $V'$. 
%We say that $J$ and $J'$ are \emph{equivalent}, written $J\iequiv J'$, if their interpretations restricted to the common variables are the same, that is, if 
%$\{\restrict I {V\cap V'} \mid I \in J\} = \{\restrict {I'} {V\cap V'} \mid I' \in J'\}$. We may write $J\eqwrt{V\cap V'} J'$ if we wish to made the common variables explicit.

The \emph{Parikh image} of a word $w\in \Sigma^*$ is the assignment 
$\parikhof w$ that for every character $a\in\Sigma$ maps the \emph{Parikh variable} $\#a$ to the number of occurrences of $a$ in $w$.
Formally, let $\pvarsof S$ denote the set of Parikh variables $\{\#s \mid s\in S\}$ obtained from a set $S$. 
The Parikh image of $w$ is a function $\parikhof w:\pvarsof \Sigma \rightarrow \nat$ such that $\parikhof w (\#a) = |w|_a$. 
The Parikh image of a language $L$ is the set of Parikh images $\parikhof L = \{\parikhof w \mid w \in  L\}$. %and the Parikh image of an FA $A$ is Parikh image of its language, $\parikhof{A} = \parikhof{L(A)}$. 
%
It is well known that $\parikhof {L(A)}$ can be efficiently computed in the form of linear formula:

\begin{lemma}[\cite{SeidlSMH04}]
$\parikhof{L(A)}$ of an FA $A$ can be computed in the form of a linear formula  $\parikhfof{A}$ such that 
%$\modelsof{\parikhfof A}_{\#\Sigma} = \parikhof{L(A)}$, 
$\modelsof{\parikhfof A} = \parikhof{L(A)}$, 
in time linear to the size of $A$.
\end{lemma}

@@

The crucial feature of parametric flat automata is that their semantics can be faithfully represented by a linear formula and handled efficiently by an SMT solver. 
\lh{, allowing to circumvent more costly automata algorithms and decision procedures. Concretize and justify this better?}
%
This is facilitated by the structural constraints of PFA, which guarantee that 
every parametric word $x\in L(A)$ is uniquely determined by its Parikh image $\parikhof{x}$, i.e., $\decode A (\parikhof x)) = x$, and analogously for the entire language of $A$, as formalized by the following lemma.

\begin{lemma}\label{lemma:parikh}
For an PFA $A$ and $x,y\in L(A)$, if $\parikhof x = \parikhof y$ then $x = y$. 
\end{lemma}
%
Indeed, since every variable appears on at most one transition, then the $\parikhof x$ value of all variables appearing within the same cycle is the same, and it is equal to the number of repetitions of that cycle in the accepting run. 
%
The accepting run on $x$ (and so $x$ itself) can thus be reconstructed from $\parikhof x$. 
%
For example, in the automaton of Figure~\ref{fig:sfa_def}, 
from $\parikhwof x {v_0} = \parikhwof x {v_1} = \cdots = \parikhwof x {v_{m-1}}=1$ and $\parikhwof x {v^0_1}=\parikhwof x {v^1_1}=\parikhwof x {v^2_1}=2$ we derive that $x=v_0(v^0_1v^1_1v^2_1)^2v_1\cdots v_{m-1}$. 
We may say that $\parikhof x$ is an encoding of $x$, and define the decoding function $\decode A$ such that $\decode A(\parikhof x) = x$, and we lift the function to sets of integer assignments in the standard manner.
%

%\begin{lemma}\label{lemma:decx}
%For a PFA $A$, $L(A) = \{\decode A(\pim) \mid \pim\in\semof{\parikhfof A}\}$.
%\end{lemma}
%\lh{should we replace $\semof{\parikhfof A}$ by $\parikhof{A}$ here and almost everywhere?}
%\yfc{Is $\parikhof{A}$ defined only over $\#I$? Does this work for Corollary~\ref{corollary:pfa}?}

\begin{lemma}\label{lemma:decx}
%For a PFA $A$, $L(A) = \{\decode A(\pim) \mid \pim\in \parikhof A\}$.
For a PFA $A$, $L(A) = \decode A(\parikhof{L(A)}$.
\end{lemma}
%\lh{should we replace $\semof{\parikhfof A}$ by $\parikhof{A}$ here and almost everywhere?}
%\yfc{Is $\parikhof{A}$ defined only over $\#I$? Does this work for Corollary~\ref{corollary:pfa}?}
%
\begin{proof}
The function $\decode A$ can particularly be implemented as follows. Given $\pim:\pvars\rightarrow\nat$,
and assuming that the lengths of the loops of $A$ are $l_0,\ldots,l_{m-1}$, 
$\decode A(\pim)$ is constructed as the word $x_0 v_{0} \ldots v_{m-1} x_{m-1}$ where for each $i\in [0,m-1]$,
$x_i = (v_i^0 \cdots v_i^{l_{i-1}})^{\pvarsof {v_i^0}}$ if $l_i >0$ and $x_i = \epsilon$ if $l_i = 0$. 
\end{proof}

Since every word $w \in \semof{A}$ is obtained from a parametric word $x\in L(A)$ and an interpretation of character variables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric word $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different words are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup \parikhof x) = w$.
The following corrolary of Lemma~\ref{lemma:decx} states that encodings of words in $\semof{A}$ are encodings of parametric words of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$:%, which are precisely all models of the formula $\parikhfof A$ interpreted over $\cvar\cup\pvarsof \cvar$, as formulated by the following corollary of Lemma~\ref{lemma:decx}:
%
%\begin{corollary}\label{corollary:pfa}
%For an PFA $A$ over character variables $V$,
%$\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}$.
%%$\semof{A} = \{I(\decode A(\pim)) \mid \pim\in\semof{\parikhfof A},I:\cvars\rightarrow\Sigma_\epsilon\}$.
%\end{corollary}

\begin{corollary}\label{corollary:pfa}
For an PFA $A$ over character variables $V$,
%$\semof{A} = \{(\decode A(I\cup\pim)) \mid \pim\in\parikhof A,I:\cvars\rightarrow\Sigma_\epsilon\}$.
$\semof A = \decode A \{(I\cup \pim)\mid \pim\in\parikhof{L(A)} ,I:\cvars\rightarrow\Sigma_\epsilon\}$
\end{corollary}
%
A minor technical difficulty is that we wish to work with pure arithmetic formulas, 
but we may also wish to specify when a character variable is assigned $\epsilon$, which is not a number. 
Hence, to allow purely arithmetic formulas to talk about the value $\epsilon$, we encode $\epsilon$ as $-1$.
In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulas constraining interpretations of character variables.

@@

The correctness of this construction is stated in Lemma~\ref{lemma:synccorrect} below. The lemma can also be understood as a precise specification of the semantics of the synchronization formula, on which we rely for correctness of the construction of under-approximations of string constraints in Sections~\ref{section:mem}, \ref{section:eq}, and \ref{section:s2i}. 
%
%\begin{lemma}\label{lemma:synccorrect}
%$
%\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} = 
%\{
%(I\cup\parikhof{x})\cup(I'\cup\parikhof{x'}) \mid 
%x\in L(A), x'\in L(A'), I(x) = I(x')
%\}
%$
%\end{lemma}

%\begin{lemma}\label{lemma:synccorrect}
%$
%\semof {\syncfof A {A'}} \iequiv
%\{
%(I\cup\parikhof{x})\cup(I'\cup\parikhof{x'}) \mid 
%x\in L(A), x'\in L(A'), I(x) = I(x')
%\}
%$
%\end{lemma}

\begin{lemma}\label{lemma:synccorrect}
$
\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} =
%\semof {\syncfof A {A'}} \iequiv
\{(I\cup I_\#)\cup (I' \cup I'_\#)\mid (I\cup I_\#) \in \encode A (w), (I' \cup I'_\#)\in\encode {A'} (w),w\in\semof A \cap \semof {A'}\} 
$
\end{lemma}

Informally, the lemma states that the models of $\syncfof A {A'}$ encode precisely the pairs of equivalent encodings of words from $\semof{A}$ and $\semof{A'}$, that constitute the intersection $\semof{A} \cap \semof{A'}$. A minor technical difficulty is that the models of $\syncfof A {A'}$ include also an assignment to the auxiliary variables of $(\cvarone_\epsilon\times\cvartwo_\epsilon) \cup \pvarsof{(\cvarone_\epsilon\times\cvartwo_\epsilon)}$. Hence in the lemma, we use $\iequiv$ instead of $=$ to restrict the models to the original variables $\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo$. %This is  achieve by quantifying out the variables of $V_\times$.

Notice that if $A$ is flat (or, symmetrically, if $A'$ is flat), then the semantic intersection $\semof A \cap \semof {A'}$ can be still decoded from the synchronisation formula $\semof {\syncfof A {A'}}$. 
Namely, due to Corollary~\ref{corollary:pfa}, we have that 
%$\semof A \cap \semof {A'} = \{\decode A(I\cup \pim) \mid (I\cup \pim) \in \semof{\syncfof A {A'}}_{\cvarone\cup \pvarone}\}$.  
$$\semof A \cap \semof {A'} = \decode A (\semof{\syncfof A {A'}}_{\cvarone\cup \pvarone})$$

@@

\subsection{Flat Domain Restriction}
Let us now give a precise definition of how we restrict the domain of string constraints using flat automata. 
A \emph{flat domain restriction} for a string contraint $\phi$ over variables $X$ is a mapping $R$ that assigns to every variable $x\in X$ a PFA $R(x)$. 
For technical reasons that will be apparent further in this section, we require that if $x\neq  y$, then $R(x)$ and $R(y)$ have disjoint sets of character variables. We will use $V_R$ to denote the union of character variables of all $R(x),x\in X$.

The semantics of a string contraint $\Phi_I$ over a set of variables $X$ restricted by $R$ is defined as 
$\semof\phi^R = \{I\in\semof \phi \mid \forall x\in X:I(x)\in\semof{R(x)}\}$. 
In the rest of this section, we will show how to build an \emph{underapproximating formula} representing the semantics $\semof{\Psi_I}^R$ of an input string constraint, that is, a formula $\Psi_O$ such that 
$\{\decode R (\semof {\Psi_O})\} = \semof{\Psi_I}_R$.

The formula $\Psi_O$ is build intuctively to the structure of $\Psi_O$.
Namely, in the followin subsections we show how to build a formula $\underf R {\phi}$ for every atomic string constraint $\phi$, and for a Boolean combination of 
string constraints we have $\underf R {\phi \land \phi'} = \underf R {\phi} \land \underf R {\phi'}$ and $\underf R {\neg\phi} = \neg\underf R{\phi}$. The underapproximation formula is then in the form $\Psi_O \defeq \underf R {\Psi_I}$.

The correctness of this construction, that is, that 
$$\decode R (\semof {\Psi_O}) = \semof{\Psi_I}^R$$ can be proved by induction to the structure of $\Psi_I$. However, for the induction step to go through, we will need a more meticulous view into the correspondence of $\phi$ and $\underf R \phi$. Particularly, we will need to
%pay attention to how $\underf R \phi$ atually encodes the models of $\phi$, and
 ensure that $\underf R \phi$ captures exactly all $R$-encodings of $\semof \phi^R$.

%
To this end, we define the following notation.
Given a PFA $A$, and a language $L\subseteq \Sigma^*$, we define the set of $A$ encodings of $L\cap \semof A$ as $\encode A (L) = \{(I\cup\pim) \mid \decode A (I\cup \pim) \in L \cap \semof{A}\}$.
%
For an interpretation $I:X\rightarrow \Sigma^*$, we define its $R$-encoding as the set $\encode R(I) = \bigcup_{x\in X} \encode A(I(x))$. We left this notation to sets of interpretation in the standard manner, i.e., $\encode R J = \bigcup_{I \in J}\encode R (I)$ for a set of interpretation $J$.
%
The needed inductive argument of the correctness of the construction can be expressed as that 
$$\semof{\underf R \phi}_{V_R\cup V_{\#R}}  = \encode R{\semof{\phi}}$$ 
We restrict need to restrict here interpretations of $\underf R \phi$ since its construction uses some additional variables.

%\subsection{Handling Regular Constraints Symbolically} \label{section:mem}
\subsection{Underapproximation of Regular Constraints} \label{section:mem}

>>>>

\newcommand{\nat}{\mathbb{N}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
%\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\mathbb{P}}(#1)} %shouldn't it be something that takes automaton as parameter and returns formula?
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}
\newcommand{\encode}[1]{\mathit{encode}_{#1}}

@@

We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the sets of natural numbers (including 0) and 
integers. For a set $A$, we use $|A|$ to denote its size. 
For $n,m\in\nat$, we write $[n,m]$ for the set of natural numbers 
%(or symbols in $\Sigma$) 
$\{k\mid n\leq k \leq m\}$. 
An \emph{alphabet} is a finite set $\Sigma$ of \emph{characters} and a \emph{word} over $\Sigma$ is a sequence $w = a_1\ldots a_n$ of characters from $\Sigma$, with $\epsilon$ denoting the \emph{empty word}. 
We use $w_1\cdot w_2$ to denote the \emph{concatenation} of words $w_1$ and $w_2$.
$\Sigma^*$ is the set of all words over $\Sigma$, $\Sigma^+ = \Sigma^*\setminus \{\epsilon\}$, $\Sigma_\epsilon = \Sigma\cup\{\epsilon\}$, and 
A \emph{language} over $\Sigma$ is a subset $L$ of $\Sigma^*$.  
%
We use $|w|$ to denote the length of $w$ and $|w|_a$ to denote the number of occurrences of the character $a\in \Sigma$ in $w$.

@@

Through the paper, we will use quantifier-free linear integer arithmetic formulas, and call them \emph{linear formulas} for short.   
Given a linear formula $\phi$, a set of variables $V$, and an \emph{interpretation} of $V$, i.e., a function $I:V\rightarrow \nat$, 
we denote by $I\models_V \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a \emph{model} of $\phi$ over $V$. \lh{Our use of indexing by $V$ in the text later is not very nice. Add this for clarity?: If $V$ contain additional variables not present in $\phi$, the these might be assigned arbitrary values by $I$. If $V$ is missing some of the variables of $\phi$, then it must be possible to complete $I$ with assignment of the missing variables into a satisfying assignment of $\phi$ (as in existential quantification).}
\lh{or really have just an equivalence of semantics indexed by variables?}
We use $\modelsof \phi$ to denote the set all models of $\phi$ over $V$. 
We drop the reference to $V$ and write just ``model of $\phi$'',  $\modelsof \phi$, and $\models$ when no confusion may arise. For a set of interpretations $S$, we use $S_\Sigma$ to denote the set of interpretations obtained from $S$ by projecting the domain to $\Sigma$.

The \emph{Parikh image} of a word $w\in \Sigma^*$ is the assignment 
$\parikhof w$ that for every character $a\in\Sigma$ maps the \emph{Parikh variable} $\#a$ to the number of occurrences of $a$ in $w$.
Formally, let $\pvarsof S$ denote the set of Parikh variables $\{\#s \mid s\in S\}$ obtained from a set $S$. 
The Parikh image of $w$ is a function $\parikhof w:\pvarsof \Sigma \rightarrow \nat$ such that $\parikhof w (\#a) = |w|_a$. 
The Parikh image of a language $L$ is the set of Parikh images $\parikhof L = \{\parikhof w \mid w \in  L\}$ and the Parikh image of an FA $A$ is Parikh image of its language, $\parikhof{A} = \parikhof{L(A)}$. 
%
It is well known that $\parikhof A$ can be efficiently computed in the form of linear formula:

\begin{lemma}[\cite{SeidlSMH04}]
$\parikhof{A}$ of an FA $A$ can be computed in the form of a linear formula  $\parikhfof{A}$ such that $\modelsof{\parikhfof A}_{\#\Sigma} = \parikhof{A}$, in time linear to the size of $A$.
\end{lemma}

@@

The crucial feature of parametric flat automata is that their semantics can be faithfully represented by a linear formula and handled efficiently by an SMT solver. 
\lh{, allowing to circumvent more costly automata algorithms and decision procedures. Concretize and justify this better?}
%
This is facilitated by the structural constraints of PFA, which guarantee that 
every parametric word $x\in L(A)$ is uniquely determined by its Parikh image $\parikhof{x}$, as formalized by the following lemma.

\begin{lemma}\label{lemma:parikh}
For an PFA $A$ and $x,y\in L(A)$, if $\parikhof x = \parikhof y$ then $x = y$. 
\end{lemma}
%
Indeed, since every variable appears on at most one transition, then the $\parikhof x$ value of all variables appearing within the same cycle is the same, and it is equal to the number of repetitions of that cycle in the accepting run. 
%
The accepting run on $x$ (and so $x$ itself) can thus be reconstructed from $\parikhof x$. 
%
For example, in the automaton of Figure~\ref{fig:sfa_def}, 
from $\parikhwof x {v_0} = \parikhwof x {v_1} = \cdots = \parikhwof x {v_{m-1}}=1$ and $\parikhwof x {v^0_1}=\parikhwof x {v^1_1}=\parikhwof x {v^2_1}=2$ we derive that $x=v_0(v^0_1v^1_1v^2_1)^2v_1\cdots v_{m-1}$. 
We may say that $\parikhof x$ is an encoding of $x$, and define the decoding function $\decode A$ such that $\decode A(\parikhof x) = x$.
%

\begin{lemma}\label{lemma:decx}
For a PFA $A$, $L(A) = \{\decode A(\pim) \mid \pim\in\semof{\parikhfof A}\}$.
\lh{should we replace $\semof{\parikhfof A}$ by $\parikhof{A}$ here and almost everywhere?}
\yfc{Is $\parikhof{A}$ defined only over $\#I$? Does this work for Corollary~\ref{corollary:pfa}?}
\end{lemma}
%
\begin{proof}
The function $\decode A$ can particularly be implemented as follows. Given $\pim:\pvars\rightarrow\nat$,
and assuming that the lengths of the loops of $A$ are $l_0,\ldots,l_{m-1}$, 
$\decode A(\pim)$ is constructed as the word $x_0 v_{0} \ldots v_{m-1} x_{m-1}$ where for each $i\in [0,m-1]$,
$x_i = (v_i^0 \cdots v_i^{l_{i-1}})^{\pvarsof {v_i^0}}$ if $l_i >0$ and $x_i = \epsilon$ if $l_i = 0$. 
\end{proof}

Since every word $w \in \semof{A}$ is obtained from a parametric word $x\in L(A)$ and an interpretation of character variables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric word $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different words are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup \parikhof x) = w$.
Encodings of words in $\semof{A}$ are encodings of parametric words of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$, which are precisely all models of the formula $\parikhfof A$ interpreted over $\cvar\cup\pvarsof \cvar$, as formulated by the following corollary of Lemma~\ref{lemma:decx}:
%
\begin{corollary}\label{corollary:pfa}
For an PFA $A$ over character variables $V$,
$\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}$.
%$\semof{A} = \{I(\decode A(\pim)) \mid \pim\in\semof{\parikhfof A},I:\cvars\rightarrow\Sigma_\epsilon\}$.
\end{corollary}
%
A minor technical difficulty is that we wish to work with pure arithmetic formulas, 
but we may also wish to specify when a character variable is assigned $\epsilon$, which is not a number. 
Hence, to allow purely arithmetic formulas to talk about the value $\epsilon$, we encode $\epsilon$ as $-1$.
In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulas constraining interpretations of character variables.

@@

The correctness of this construction is stated in Lemma~\ref{lemma:synccorrect} below. The lemma can also be understood as a precise specification of the semantics of the synchronization formula, on which we rely for correctness of the construction of under-approximations of string constraints in Sections~\ref{section:mem}, \ref{section:eq}, and \ref{section:s2i}. 
%
\begin{lemma}\label{lemma:synccorrect}
$
\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} = 
\{
(I\cup\parikhof{x})\cup(I'\cup\parikhof{x'}) \mid 
x\in L(A), x'\in L(A'), I(x) = I(x')
\}
$
\end{lemma}
Informally, the lemma states that the models of $\syncfof A {A'}$ encode precisely the pairs of equivalent encodings of words from $\semof{A}$ and $\semof{A'}$, that constitute the intersection $\semof{A} \cap \semof{A'}$. A minor technical difficulty is that the models of $\syncfof A {A'}$ include also an assignment to the auxiliary variables of $(\cvarone_\epsilon\times\cvartwo_\epsilon) \cup \pvarsof{(\cvarone_\epsilon\times\cvartwo_\epsilon)}$. Hence in the lemma, we have to restrict the models to the original variables $\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo$. %This is  achieve by quantifying out the variables of $V_\times$.

Notice that if $A$ is flat (or, symmetrically, if $A'$ is flat), then the semantic intersection $\semof A \cap \semof {A'}$ can be still decoded from the synchronisation formula $\semof {\syncfof A {A'}}$. 
Namely, due to Corollary~\ref{corollary:pfa}, we have that $\semof A \cap \semof {A'} = \{\decode A(I\cup \pim) \mid (I\cup \pim) \in \semof{\syncfof A {A'}}_{\cvarone\cup \pvarone}\}$.

@@

\subsection{Flat Domain Restriction}

Given a PFA $A$, and a language $L\subseteq \Sigma^*$, we define the set of $A$ encodings of $L\cap \semof A$ as
the set $\encode A (L) = \{(I\cup\pim) \mid \decode A (I\cup \pim) \in L \cap \semof{A}\}$.
%
A domain restriction $R$ for a string contraint $\phi$ over variables $X$ is a mapping that assigns to every variable $x\in X$ a PFA $R(x)$ such that for different variables $x,y \in X$, $R(x)$ and $R(y)$ have disjoint sets of variables. 
%
For an interpretation $I:X\rightarrow \Sigma^*$, we define its $R$-encoding as the set $\encode R(I) = \bigcup_{x\in X} \encode A(I(x))$, and we left this notation to sets of interpretation in the standard manner, i.e. $\encode R J = \bigcup_{I \in J}\encode R (I)$.
%
In the following section, we will discuss how, given a string contraint $\phi$ and a domain restriction $R$,
to build an arithmetic formula $\Psi$ such that $\semof\Psi_{V_R\cup\#V_R} = \encode R(\semof{\phi})$. 
Tha is, formula $\Psi$ will capture $R$-encodings of soltuions of $\phi$ under the domain restriction given by $R$.

\subsection{Handling Regular Constraints Symbolically} \label{section:mem}

Assume that the constraint we take from $\Phi_I$ is a regular constraint $\phi_r \defeq x \in L(A)$. 
%Assume that $A=(Q,T,\Sigma,q_i,q_f)$. 
We represent the FA $A$ by a pair of a PA $A'$ and linear formula $\Psi_{char}$. 
The PA $A'$ operates over fresh character variables $v_a, a\in\Sigma_\epsilon$, and is obtained from $A$ by replacing every occurrence of each character $a\in\Sigma_{\epsilon}$ on a transition by the variable $v_a$. 
The formula $\Psi_{char}$ then binds the fresh character variables to the character values they represent, 
namely, $\Psi_{char} = \bigwedge_{a\in\Sigma_\epsilon} v_a = a$.
The resulting under-approximation of $\Psi_r$ is then the formula $\syncfof{A_x} {A'} \land \Psi_{char}$, which will be added to $\Phi_O$.

The following lemma states the correctness of this construction, and it follows from Corollary~\ref{corollary:pfa}, Lemma~\ref{lemma:synccorrect}. 
Intuitively, it says that the resulting constraint represents precisely the restriction of $L(A)$ to the domain $\semof{A_x}$, the $A_x$-encodings of the strings in $L(A)\cap\semof{A_x}$. 
%
\begin{lemma}\label{lemma:memcorrect}  ${\semof {\syncfof {A_x }{A'} \land \Psi_{char}} } _{V_x \cup \pvarsof{V_x}} = \{(I\cup\pim) \mid \decode {A_x}(I\cup\pim) \in L(A)\}$. 
\end{lemma}
%
(An alternative description)
\begin{lemma}\label{lemma:memcorrect} For all $w\in \semof{A_x}$, %$(I\cup\pim)$ in ${\semof {\syncfof {A_x }{A'} \land \Psi_{char}} } _{V_x \cup \pvarsof{V_x}}$ such that $w =\decode {A_x}(I\cup\pim)$ if and only if $w \in L(A)$.
%	$$\exists (I\cup\pim) \in {\semof {\syncfof {A_x }{A'} \land \Psi_{char}} } _{V_x \cup \pvarsof{V_x}}. w =\decode {A_x}(I\cup\pim)$$
	\begin{center}
	$\exists (I\cup\pim)$ in ${\semof {\syncfof {A_x }{A'} \land \Psi_{char}} } _{V_x \cup \pvarsof{V_x}}$ s.t. $w =\decode {A_x}(I\cup\pim)$\\
		iff\\
		$w \in L(A)$.
	\end{center}
%	${\semof {\syncfof {A_x }{A'} \land \Psi_{char}} } _{V_x \cup \pvarsof{V_x}} = \{(I\cup\pim) \mid \decode {A_x}(I\cup\pim) \in L(A)\}$. 
\end{lemma}

\lh{
\begin{lemma}
${\semof {\syncfof {A_x }{A'} \land \Psi_{char}} } _{V_x \cup \pvarsof{V_x}} = \encode {A_x}(L(A))$
\end{lemma}
}
